---
title: "Workflow of Final Project"
output: html_notebook
---


###Part I. Clustering

**1. Set up**
```{r}
library(lazyeval)
library(ggplot2)
library(factoextra)
cluster_table <- read.table("table_clusters.txt", header = T) # import file from comparative analysis table
cluster_table <- as.data.frame(cluster_table)
```

**2. Sort table based on total number of clusters across species**
```{r}
cluster_table$sum <- rowSums(cluster_table)
cluster_table <- cluster_table[order(-cluster_table$sum),]
cluster_table <- subset(cluster_table, select = -c(sum))
cluster_table <- head(cluster_table,1000)
```

**3. Creating table of A clusters and D clusters (check final table, there was an A2 not in line with the rest of "A"s)**
```{r}
Acluster_table <- cluster_table[,1:8] #creating table of A clusters
Dcluster_table <- cluster_table[,9:13] #creating table of D clusters
```

**4. No normalization PCA??**
```{r}
cluster_table <- t(cluster_table)
cluster.pca <- prcomp(cluster_table, scale = TRUE, na.rm = TRUE)
ind.coord <- cluster.pca$x
```

**5. Subsetting data, taking first 1000 columns**
```{r}
subsections <- as.data.frame(cluster_table[,1:2])
subsections$sub <-c("A1", "A1", "A1", "A2", "A2", "A2", "A2", "D5", "D5", "D5", "D5", "A2", "D5")

subsections <- subsections[,-c(1:2)]
subfac <- as.factor(subsections)

png("cotton_outgroup.PCA.direct.annot.new.png", 1000, 1000, pointsize=20)
fviz_pca_ind(cluster.pca, habillage=subfac) + theme_minimal()
dev.off()
```


###Part II. Recreate the bar plot showing the abundance of the most common repeat types in the genomes of three *Gossypium* species (Figure 2)

**1. Set up**
```{r}
library(ggplot2)
library(reshape2)
stable <- read.table("f.csv", sep = ",", header =TRUE) #Read in data frame
```

**2. Create new rows that are averages of each species**
```{r}
stable$A1 <- rowMeans(stable[,2:4])
stable$A2 <- rowMeans(stable[,5:9])
stable$D5 <- rowMeans(stable[,10:14])
```

**3. Create standard deviation function**
```{r}
std <- function(x) sd(x)/sqrt(length(x))
```

**4. Find standard deviation of each species**
```{r}
stable$A1err <- apply(stable[,2:4], 1, std)
stable$A2err <- apply(stable[,5:9], 1, std)
stable$D5err <- apply(stable[,10:14], 1, std)
```

**5. Create Standard Deviation minimum**
```{r}
stable$A1min <- stable$A1 - stable$A1err
stable$A2min <- stable$A2 - stable$A2err
stable$D5min <- stable$D5 - stable$D5err
```

**6. Create Standard Deviation maximum**
```{r}
stable$A1max <- stable$A1 + stable$A1err
stable$A2max <- stable$A2 + stable$A2err
stable$D5max <- stable$D5 + stable$D5err
```

**7. Remove categorizations of TE that we don't care about**
```{r}
stable <- stable[,-(2:14)]
```

**8. Melt dataframe in order to plot properly**
```{r}
melted <- melt(stable[,(1:4)])
min <- c(stable$A1min, stable$A2min, stable$D5min)
max <- c(stable$A1max, stable$A2max, stable$D5max)
melted$min <- min
melted$max <- max
```

**9. Create standard error bars**
```{r}
limits <- aes(ymax=melted$max, ymin=melted$min)
```

**10. Maintain width of bars**
```{r}
dodge <- position_dodge(width=0.9)
```

**11. Create image space**
```{r}
png("Figure_TE.amounts.png", 7500, 5000, pointsize=12, res=600)
```

**12. Plot within image space**
```{r}
ggplot(melted, aes(x=Lineage, y=value, fill = variable)) + geom_bar(stat = "identity",position = dodge) + geom_errorbar(limits, position = dodge, width=0.3) + labs(y="Mbp/1C", x="") + theme_set(theme_grey(base_size=12)) + scale_fill_hue(l=40)
dev.off()
```


###Part III. Recreate the two-way ANOVA of cluster abundance among the largest 1000 clusters in four *Gossypium* species (Table 2)

**1. Set up**
```{r}
library(reshape2)
new_table = cluster_table
row.names(new_table) <- c("A1", "A1", "A1", "A2", "A2", "A2", "A2", "D5", "D5", "D5", "D5", "A2", "D5")
```

**2. Melt dataframe in order to plot properly**
```{r}
melted <- melt(new_table)
Y <- melted$value
Species <- as.factor(melted$Var1)
Cluster <- as.factor(melted$Var2)
```

**3. ANOVA**
```{r}
anova(lm(Y~Cluster*Species))
```


###Part IV. Recreate scatter plot of cluster abundance in *Gossypium herbaceum*(A1) and *Gossypium arboreum*(A2) (Figure 3)

**1. Set up**
```{r}
library(ggplot2)
new_table <- as.data.frame(t(cluster_table))
new_table$A1ave <- rowMeans(subset(new_table, select = c(A1_155_, A1_073_, A1_097_), na.rm = TRUE))
new_table$A2ave <- rowMeans(subset(new_table, select = c(A2_255_, A2_034_, A2_044_, A2_099_, A2_101_), na.rm = TRUE))
```

**2. Create the scatter plot**
```{r}
ggplot(A1ave~A2ave, data=new_table, mapping = aes(x = A1ave, y = A2ave)) + geom_point() + geom_abline(intercept=0, slope=1)

png("cotton.ordination.png", 5000, 5000, pointsize=12, res=600)
ggplot(cmdpoints, aes(x=V1, y=V2, color=genome, shape=genome)) + geom_point(size=2) + xlab("PCoA component 1") + ylab("PCoA component 2") + scale_color_manual(values=c("A1"="orchid", "A2"="orchid4", "D5"="slategrey","kirkii"="blue3", "kokia"="green3"))+ geom_text_repel(aes(label=species))### Recreate Figure 4 Simon's way
mydata <- t(annot_clust[,-c(1)])
d <- dist(mydata, method = "euclidean")

cmdfit <- cmdscale(d,eig=TRUE, k=2) # k is the number of dim

x <- cmdfit$points[,1]
y <- cmdfit$points[,2]
plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2", 
     main="Metric        MDS",        type="n")
text(x, y, labels = row.names(mydata), cex=.7)
```


###Part V. Recreate the plot showing cotton samples grouped by repeat content (Figure 4)

**1. Make the euclidean distance matrix**
```{r}
d <- dist(cluster_table, method = "euclidean")
```

**2. Using cmdscale to perform classical MDS (principal coordinate analysis)**
```{r}
cmdfit <- cmdscale(d,eig=TRUE, k=2) # k is the number of dim
x <- cmdfit$points[,1]
y <- cmdfit$points[,2]
```

**3. Make the plot**
```{r}
png("SRB.Fig4.png", 5000, 5000, pointsize=12, res=600)
plot(x, y, xlab="Coordinate 1", ylab="Coordinate 2", main="Metric MDS",	type="n")
text(x, y, labels = row.names(cluster_table), cex=.7)
points(x, y, pch=19)
dev.off()
```